---
title: "The movie industry has been rejecting single labels since the 1950s"
author: "Emma Livingston"
date: "December 1, 2017"
output:
  html_document:
    code_folding: hide
 #   theme: make it look different
---


```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mdsr)
library(RMySQL)
library(tidyverse)
db <- dbConnect_scidb(dbname = "imdb")
library(fivethirtyeight)
library(plotly)
library(RColorBrewer)
```

Have films gotten more greedy in their pursuit of an audience? By sorting a movie into genres, production companies are better able to attract viewers who know they will enjoy the movie and dissuade viewers who may not be interested. However, if they're selecting more genres, they're doing more of the former and less of the latter. We looked at the number of genres attached to movies over time to find out.

Film producers know that not everyone likes every genre, and they don't want to alienate potential new viewers by only having their movie fit in one box. For example, your typical superhero action movie has a romantic subplot, some comic relief, and a sci-fi or fantasy origin story. However, there's a difference between it being called an action movie and it being called a sci-fi/fantasy action rom-com. The former sounds fairly typical, and the latter sounds both like it has something for everyone and like it's a bit confused. But it might attract a wider audience by being firmly placed into all these different genres. And over time, producers have been learning this, since the average Hollywood movie has been placed into an increasing number of genres since the 1990s.

We started with data from the [IMDb database](/http://www.imdb.com/), which relies on user submissions that go through an [approval process](/https://help.imdb.com/article/contribution/contribution-information/when-will-my-update-be-added/GEKS9SBHDEHBAVC7?ref_=helpms_helpart_inline#).

```{r}
num_genres <- db %>%
  dbGetQuery("SELECT title.title, title.id, production_year, mi1.info AS genre, mi2.info AS country, sum(1) AS N
FROM title
JOIN movie_info AS mi1 ON title.id = mi1.movie_id
JOIN movie_info AS mi2 ON title.id = mi2.movie_id
WHERE mi1.info_type_id = 3 AND kind_id = 1 
AND mi2.info_type_id = 8 AND mi2.info = 'USA'
GROUP BY title.id
ORDER BY title, title.id;")
#This query makes use of the index on info_type_id in the movie_info table
#Group by movie to find the number of genres per movie

num_year <- db %>%
  dbGetQuery("SELECT production_year, info, sum(1) AS N
FROM movie_info
JOIN title ON title.id = movie_info.movie_id
WHERE info_type_id = 3 AND kind_id = 1
GROUP BY production_year
ORDER BY title, title.id;")
#This query also uses the index on the info_type_id in the movie_info table
#Group by production year to find the number of movies per year

joined <- num_genres %>%
  inner_join(num_year, by = "production_year") %>%
  filter(production_year <= 2026) %>%
  group_by(production_year) %>%
  summarize(avg_genres = mean(N.x))
#Here we combined the resulting tables from the above queries, filtered out any film with a production year greater than 2026 (because of the \href{http://www.imdb.com/title/tt5174640/}{film to be released in 2115}), and found the average number of genres per year.

past_joined <- joined %>%
  filter(production_year <= 2017)
```

By combining that data with a [dataset](/https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/)[^1] that has information on the profits each movie made, we were able to come up with a plot that included both the profit and the number of genres. 

```{r}
joined_bechdel <- num_genres %>%
  right_join(bechdel, by = "title") %>%
  filter(production_year == year) %>%
  mutate(Profit = domgross_2013 - budget_2013) %>%
  rename(Genres = N)
profit <- ggplot(data = joined_bechdel, aes(x = year, y = Profit, color = Genres, label = title)) +
  geom_smooth(size = 1.2, col = "black") +
  geom_point(alpha = 0.5) +
  scale_color_distiller(palette = "YlGnBu", direction = "-YlGnBu") +
  scale_y_continuous(breaks = seq(from = -500000000, to = 1500000000, by = 500000000), labels = c("-0.5", "0", "0.5", "1.0", "1.5")) + 
 xlab("Year") +
 ylab("Profit in 2013 Dollars (billions)") +
  theme(panel.background = element_rect(fill = "dark grey")) 
ggplotly(profit)
```

While [box office](/https://www.the-numbers.com/market/) revenues are increasing, the number of ticket sales has been diving since the early 2000s, and the profits have leveled off as well. So if you are a movie producer in this seemingly bleak landscape, you have probably started looking to past successes of Hollywood for inspiration on how to pad your pockets. (INSERT SOME STUFF HERE ABOUT THE GOLDEN AGE OF HOLLYWOOD).

```{r}
ggplot(data = past_joined, aes(x = production_year, y = avg_genres)) +
  geom_line(size = 1.2) +
  geom_point(aes(x = 2017, y = 2.077276), shape = 20, size = 5, col = "cornflowerblue") +
  geom_text(aes(label = "2017", x = 2012, y = 1.75), col = "cornflowerblue") +
  geom_point(aes(x = 1947, y = 2.473159), shape = 20, size = 5, col = "cornflowerblue") +
  geom_text(aes(label = "1947", x = 1942, y = 2.25), col = "cornflowerblue") +
  ylab("Average Genres per Movie") +
  xlab("Production Year") +
  scale_x_continuous("", breaks = seq(from = 1880, to = 2020, by = 20))
``` 

(SOME GOOD EXPLANATION OF THIS GRAPH)

Some of the big changes on the graph aren't that surprising when put into historical context. For example, take the steady decline of average genres per movie in the late 1950s and 60s. This is when television became an important form of entertainment, and it uses the same kinds of resources as movies do. With a boom in TV, it makes sense that movies would be less complex for a while.

The recent rise in genres per movie can also be explained by historical trends as well. The last steady rise in this number was from the 1920s and 30s, which could be called the golden age of cinema, with new innovations such as sound and color. With movie profits being far from on the rise, Hollywood producers could be trying to imitate this success.

So the number of genre labels each movie gets has been growing overall. And looking at the next few years, that trend is predicted to skyrocket, according to the movies already entered in IMDb.[^2]

[^2] Any movies that haven't come out yet that are already on IMDb are likely sequels that already have release dates. These data are might not be representative of the trends that we will see once those dates arrive, but it is valuable to see where production studios believe they are headed.


```{r}
future <- joined %>%
  filter(production_year > 2017)
print(future)
```

So from a marketing point-of-view, it makes sense to try to fit your movie into as many genres as you can to draw in the biggest audience possible to up those ticket sales. More genres equals more money, right? Not necessarily.


[^1] This data comes from a combination of [bechdeltest.com](/http://bechdeltest.com/) and [www.the-numbers.com/](/https://www.the-numbers.com/), which we have access to through a previous project. This data is only inclusive of movies on both websites, meaning that it is not all encompassing. On top of that, the only movies in the graphs involving profit must be in both the `bechdel` dataset and the IMDb database. 



```{r}
ggplot(joined_bechdel, aes(x = Genres, y = Profit)) +
  geom_jitter() +
  scale_y_continuous(breaks = seq(from = -500000000, to = 1500000000, by = 500000000), labels = c("-0.5", "0", "0.5", "1.0", "1.5")) +
  ylab("Profit in 2013 Dollars (billions)")
```

If it were true that more genres mean more profit, we would expect to see a much higher profit from the movies that have more genres. Instead, what we see is a peak of profit in the 2-4 genre range. That means that a movie listed as [comedy, romance, and drama](/http://www.imdb.com/title/tt0259446/?ref_=ttls_li_tt) is likely to do better than a movie listed as an [action-comedy-crime-fantasy-horror-mystery-Sci-Fi-thriller](/http://www.imdb.com/title/tt1013860/). 


